Deep Learning for Automatic Speech Recognition

### 2015-04-21 Paper/Paper.tex
Presentation slides
Added in my rough cut of what I plan on talking about Thursday.
Ryan's initial update

\paragraph{} In this paper we examine two deep learning method for automatic speech recognition: deep neural networks and convolutional neural networks based on the works of \cite{DBLP:journals/taslp/DahlYDA12} and  \cite{DBLP:journals/taslp/Abdel-HamidMJDPY14} respectively. \TODO{Ryan TODO}

\section*{Traditional Approach} \TODO{Ryan TODO}

@@

\begin{equation}
E(v, h) - b^T v - c^T h - v^T W h
\end{equation}

@@

>>>>

\paragraph{} In this paper we examine two deep learning method for automatic speech recognition: deep neural networks and convolutional neural networks based on the works of \cite{DBLP:journals/taslp/DahlYDA12} and  \cite{DBLP:journals/taslp/Abdel-HamidMJDPY14} respectively. \TODO{Ryan - Talk about speech recognition in general, linguistics variations, }

\section*{Traditional Approach} \TODO{Ryan GMM-HMMs}

\paragraph{}

@@

\begin{equation}
E(v, h) = - b^T v - c^T h - v^T W h
\end{equation}

@@

\paragraph{} In this section we explore the work done in \cite{DBLP:journals/taslp/Abdel-HamidMJDPY14} on the use of convolutional neural network-hidden markov model hybrid (CNN-HMM) for phone recognition performed on the TIMIT dataset and a large-vocabulary voice search task. In this model, temporal variability is handled by the HMM, while convolutions occur along the frequency axis of the acoustic signal. We will begin by exploring the overall structure of the network, addressing the shape of the input and looking closely at the convolution and pooling plys. Pre-training is similar to that used in \cite{DBLP:journals/taslp/DahlYDA12}, so we will instead focus on the fine-tuning of  parameters using the back-propagation algorithm. We will then transition into the results section where we will compare the performance of the CD-DNN-HMM with that of the CNN-HMM.

\TODO{Go over structure of input, why not MFCCs, why frequency axis}

\paragraph{} Over the past few years, convolutional neural networks have proven to be highly successful in image recognition tasks. Not only have they achieved the highest results on the MNIST character recognition dataset\TODO{cite}, but they've even proven to outstrip hand-crafted object recognition systems on the much more complex ImageNet dataset.

\subsection*{Convolution Ply}

\paragraph{}Without padding the input with dummy values, the convolution operation would produce a lower-dimensional feature map where each dimension decreases by the filter size \textit{F} minus one. If this is done, there will be the same number of units in the input feature map as in that of the convolution ply.

\subsection*{Pooling Ply}

\paragraph{}The pooling ply in the network acts as a next step from the convolution ply where the resolution of the feature maps is reduced. \TODO{Why is this important}

\paragraph{}The pooling function used in a CNN is normally either one of \textit{averaging} or \textit{maximization}. In this work, a max-pooling function was used throughout the network:
\begin{equation*}
p_{i,m} = \max_{n=1}^{G} q_{i, (m-1)xs+n}
\end{equation*}
where \textit{G} is the pooling size and \textit{s} is the shift size. If $G=s$, there is no overlap between pooling windows. \TODO{cite for maxpool vs average + 1 sent} \cite{DBLP:journals/taslp/Abdel-HamidMJDPY14} decided to vary these independently to discover whether or not this held true for acoustic processing as well.

\subsection*{Limited Weight Sharing}
\TODO{Explain why limited compared to image processing method}

\subsection*{Training}
\TODO{back prop with weight matrices}

\paragraph{}Pre-training of the CNN is also done with RBMs, with some slight modifications to account for both convolution and pooling plies. This results in a CRBM, which will not be covered here.

### 2015-04-21 Presentation/Presentation.tex
Ryan's initial update

	\begin{enumerate}
		\item \textcolor{blue}{Introduction}
		\item Gaussian Mixture Models
		\item Convolutional Neural Networks
		\item Deep Neural Networks
		\item Experimental Results
		\item Conclusions
	\end{enumerate}
\end{frame}

\begin{frame}

\begin{frame}

\begin{frame}

\begin{frame}

@@

\begin{frame}

@@

\begin{frame}

@@

\begin{frame}

\begin{frame}

@@

\begin{frame}

@@

\begin{frame}

@@

\begin{frame}

@@

\begin{frame}

>>>>

	\begin{enumerate}
		\item \textcolor{blue}{Introduction}
		\item Gaussian Mixture Models
		\item Convolutional Neural Networks
		\item Deep Neural Networks
		\item Experimental Results
		\item Conclusions
	\end{enumerate}
\end{frame}
\begin{frame}
	\frametitle{Speech Recognition}
	\begin{center}
		Main goal of any automatic speech recognition system is to take some acoustic signal as input and convert it into a corresponding lexical representation.
	\end{center}

	\begin{center}
		Tasks range from very easy to extremely difficult: Number recognition, phone recognition, large vocabulary speech recognition, human-computer, human-human
	\end{center}

	\begin{center}
		\textbf{Many variable factors to account for}: environmental noise, speaker variation, vocabulary size, degree of fluency, etc.
	\end{center}
\end{frame}
\begin{frame}
	\frametitle{The ASR Problems}
	\begin{center}
		There are \textbf{four} substantial problems that must be solved in any statistical ASR system:
	\end{center}

	\begin{enumerate}
		\item Acoustic Processing
		\item \textcolor{blue}{Acoustic Modeling}
		\item Language Modeling
		\item The Search Problem
	\end{enumerate}
\end{frame}

\begin{frame}
	\frametitle{Acoustic Processing}
	\begin{center}
		Represent acoustic data in a way that has few model parameters, but maintains necessary linguistic information.
	\end{center}
	\begin{center}
		Main form of feature representation are the mel-frequency cepstral coefficients (MFCCs) extracted from a waveform.
	\end{center}

	\begin{itemize}
		\item Convert the spectrum of a window into its cepstrum
		\item Effectively separates the source from filter
		\item Cepstral coefficients are uncorrelated
		\item Use the first 12 (represent the filter), their deltas and delta-deltas for 39 features per frame
	\end{itemize}

	\begin{equation*}
		c[n] = \sum_{n=0}^{N-1}\log(|\sum_{n=0}^{N-1}x[n]e^{-j\frac{2\pi}{N}kn}|)e^{j\frac{2\pi}{N}kn}
	\end{equation*}
\end{frame}

\begin{frame}
	\frametitle{Acoustic Modeling}
	\begin{center}
		Given the observation feature vector, estimate an HMM state posterior probability: \begin{equation*}
		\cprob{x}{w}\end{equation*} Where \textit{x} is some acoustic signal as input and \textit{w} is a word, phone, sub-phone, diphone, or triphone. 
	\end{center}

	We will explore three ways of estimating this probability:
	\begin{itemize}
		\item Gaussian Mixture Models (GMMs)
		\item Convolutional Neural Networks (CNNs)
		\item Deep Neural Networks (DNNs)
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Language Modeling and Search}
	\begin{center}
		Job of the language model is to estimate the prior probability:
		$\prob{w}$
	\end{center}
	\begin{center}
		This is usually done with an N-gram model like we've seen throughout the semester.
	\end{center}
	\vfill
	\begin{center}
		We normally use GMM-HMM or ANN-HMM hybrids for acoustic/language models. Given these models, we need to find the best sequence of words for the acoustic input.
	\end{center}
	\begin{center}
		Search is normally performed using Viterbi, A* decoding, or N-best.
	\end{center}

@@

\begin{frame}
	\frametitle{Gaussian Mixture Models}\TODO{clean up}
	\begin{center}
		\includegraphics[scale=.3, clip=true, trim=5cm 9cm 4cm 10cm]{gmm.pdf}
	\end{center}

	\begin{center}
		Before the rise of Deep Learning approaches, GMMs were ubiquitous for representing the acoustic model.
	\end{center}
	\begin{center}
		Based on assumption that a mixture of Gaussians can be used to represent the spectral shape.
	\end{center}

@@

\begin{frame}
	\frametitle{GMM Models}
	\begin{center}
		Use a weighted mixture of multivariate Gaussians as a model to assign a likelihood to an acoustic observation.
	\end{center}

	\begin{center}
	Below is the output likelihood function $b_{j}(o_{t})$:
	\end{center}

	\begin{equation*}
	b_{j}(o_{t}) = \sum_{m=1}^{M}c_{jm}\frac{1}{\sqrt{2\pi|\sigma_{jm}|}}\exp[(x-\mu_{jm})^T\sigma_{jm}^{-1}(o_{t}-\mu_{jm})]
	\end{equation*}

	\begin{center}
		A GMM model can be trained using the Baum-Welch algorithm to iteratively recompute the mean, mixture weight, and covariance.
	\end{center}

@@

\begin{frame}
	\frametitle{The Rise of Deep Learning}
	\begin{center}
		Recently, Artificial Neural Networks (ANNs) have experienced great results in a variety of tasks, and ANN-HMMs have begun to replace GMM-HMMs.
	\end{center}
	\begin{itemize}
		\item Increasing processing power and more efficient algorithms for training.
		\item GMMs have much fewer model parameters, and are much easier to train.
		\item ANN-HMM hybrids are much larger, and can take in multiple frames (9-15) 
		\item GMM-HMMs typically only look at a single frame.
	\end{itemize}

	\begin{center}
		This increase in size allows hybrid ANN-HMMs to model highly correlated features across wide temporal contexts.
	\end{center}

@@

\begin{frame}
	\frametitle{Convolutional Neural Networks}
	\begin{center}
		\textbf{Convolutional Neural Networks for Speech Recognition}
	\end{center}

	\hfill Ossama Abdel-Hamid \hfill Abdel-rahman Mohamed \hfill Hui Jiang

	\hfill Li Deng \hfill Gerald Penn \hfill Dong Yu \hfill

	\begin{center}
		IEEE/ACM Transactions on Audio, Speech, and Language processing, Vol. 22, No. 10, October 2014
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{CNN Background}
	\begin{center}
		CNNs have achieved high levels of success in image recognition tasks. They have a very complex architecture that is influenced by our own visual cortex.
	\end{center}
		\begin{itemize}
			\item Special layer with a convolution ply and pooling ply (simple cells and complex cells)
			\item Weights are shared across all neurons in the same convolution ply
			\item Replicates features across the image in order to recognize patterns anywhere within it
		\end{itemize}
\end{frame}
\begin{frame}
	\frametitle{CNN-HMM for ASR}
	\begin{center}
		Apply a similar architecture to the task of ASR with some differences to help exhibit invariance to small perturbations in frequency caused by different speakers, emotional status, background noise, etc.:
		\vfill
		\begin{itemize}
			\item Local Connectivity: Mel-frequency spectral coefficients (MFSC) features
			\item Convolutions along frequency axis: HMMs handle time axis, allow for small frequency shifts
			\item Limited weight sharing: Same pool = Same weights
		\end{itemize}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Input Structure of the CNN}
	\includegraphics[scale=0.6, trim=1.1cm 0 0 7.5cm, clip=true]{cnn-input.pdf}

@@

\begin{frame}
	\frametitle{Global Architecture}
	\includegraphics[scale=0.65, trim=2cm 0 0 8cm, clip=true]{cnn-global.pdf}
\end{frame}

\begin{frame}
	\frametitle{Convolution Ply}
	\begin{center}
		The convolution ply introduces locality into the network. Unlike in image processing, weights are only shared between neurons that go to the same pooling ply. 
	\end{center}

	\begin{center}
		Limited weight sharing is used to reflect the fact that properties of the speech signal vary over different frequency bands. Each frequency band has its own shared weights.
	\end{center}	
	\vfill
	\begin{center}
		Each unit in a single feature map can be computed as below:
	\end{center}
	\begin{equation*}
	q_{j,m} = \sigma(\sum_{i=1}^{I}\sum_{n=1}^{F}o_{i,n+m-1}w_{i,j,n}+w_{0,j}), (j = 1, ..., J)
	\end{equation*}

@@

\begin{frame}
	\frametitle{Pooling Ply}
	\begin{center}
		The pooling ply reduces the resolution of the feature maps, serving as a generalization over the features of the convolution ply. This helps with handling small shifts in frequency.
	\end{center}
	\begin{center}
		In this paper they use a simple max-pooling function:
	\end{center}
	\begin{equation*}
	p_{i,m} = \max_{n=1}^{G} q_{i, (m-1)xs+n}
	\end{equation*}
	\begin{center}
		where \textit{G} is the pooling size and \textit{s} is the shift size. If \textit{G = s}, there is no overlap between the pooling windows.
	\end{center}

@@

\begin{frame}
	\frametitle{A Simpler Example}
	\includegraphics[scale=1, clip=true, trim=5cm 0 0 10cm]{conv-pool.pdf}
\end{frame}

\begin{frame}
	\frametitle{Training with Back-Propagation}
	\begin{center}
		Most frequently used training algorithm for ANNs. We want to discover how the output of the ANN changes with respect to the input. Both networks use a softmax output layer to compute the posterior probability:
		\begin{equation*}
		y_{i} = \frac{\exp(o_{i}^{(L)})}{\sum_{j}\exp(o_{j}^{(L)})}
		\end{equation*}
		Goal is to minimize the cross-entropy function: $Q({W^{(l)}}) = -\sum_{i}d_{i}\log y_{i}$ where \textit{d} is the target and \textit{y} is the prediction.
		\vfill
		Compute derivative of Q with respect to each weight matrix, and use stochastic gradient descent to minimize the objective function.
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Back-Propagation cont.}
	\includegraphics[scale=0.65, trim=2cm 0 0 8cm, clip=true]{backprop.pdf}
\end{frame}
